---
title: 'Final Project: Identifying Determinants of Political Corruption in Malaysia'
author: "Alif Zabidi"
date: "10/4/2020"
output:
  pdf_document:
    toc: true
    number_sections: true
bibliography: govt670_FP_references.bib
fontsize: 12pt
geometry: margin=1in
citation_package: bibtex
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r, echo=FALSE, message=FALSE, warning=FALSE}
library(tidyverse)
library(ggplot2)
library(bibtex)
library(tree)
library(dplyr)
library(leaps)
library(changepoint)
library(car)
library(randomForest)
library(caret)
```


## An Analysis of Political Corruption in Malaysia: Identification of Key Determinants of Corruption

# Introduction

According to Transparency International, Malaysia has often occupied the position of the second least corrupt nation in Southeast Asia. But corruption and the public perception of such corruption has become a major political issue in Malaysia in recent years, a result of public outcry and the steady growth of opposition to Malaysia's long-running former ruling coalition, Barisan Nasional (or in English, the National Front). Barisan Nasional was the longest running ruling coalition in Malaysia's history, having administered the country since its independence in 1957. It began as a popular coalition made up of racially oriented political parties including UMNO (the United Malays National Organisation), the MIC (Malaysian Indian Congress) and the MCA (Malaysian Chinese Association) amongst others, and was for many decades relatively unchallenged in its dominance in Malaysia's parliament [@berman2016public, pg 145-147]. However, in the past decade, Malaysia faced a tumultuous period under the government of the former Prime Minister Najib Razak, with multiple scandals and allegations of graft and other forms of corruption being leveled against the Prime Minister and Barisan Nasional [@edwards2018malaysia, pg 11-12]. One of the largest of these scandals relates to the 1 Malaysia Development Fund, often abbreviated as 1MDB, in which a national development fund was spearheaded by Najib Razak and was used to siphon massive quantities of national funding into the personal wealth of politicians and other elites within the ruling coalition [@edwards2018malaysia, pg 9-10]. 

Najib Razak had gained power in 2009, and ruled the country for nearly a decade, during which time increasing evidence of extensive government corruption came to light as a result of media coverage and investigations by both domestic and international monitoring agencies. Public outcry continued to grow to the point that it gave Barisan Nasional's political opponents sufficient momentum to mount a credible threat to BN's mandate. This came to fruition in May of 2018 during Malaysia's 14th General Election, when the country's major opposition coalition Pakatan Harapan (the Alliance of Hope), managed to overturn Barisan Nasional's long-running majority in parliament, and became the defacto ruling coalition of Malaysia. It is of interest to note that Pakatan Harapan was spearheaded by former Prime Minister Tun Mahathir Mohamad, who was Malaysia's 4th and longest reigning democratically elected leader, during his term in office between 1981 and 2003. Prime Minister Mahathir, though largely popular in his time in office, was also no stranger to allegations of corruption [@edwards2018malaysia, pg 9]. The most recent change, which occured on the 1st of March 2020 was essentially a political coup de grace, in which the Deputy Prime Minister Muhyiddin Yassin was able to secure enough support from political parties within and without the Pakatan Harapan coalition, and wrest sufficient parliamentary seats away from PM Tun Mahathir Mohamad. Due to how recent this shift in leadership was, it suffices to say that there is yet to be sufficient data for this to be included in the intended analysis. 

Given the tumultous political situation Malaysia now finds itself in, it is imperative that research should be dedicated to political corruption in Malaysia and how greatly it's effect has differed across the various governments that have ruled the country. It is the intention of this paper to identify key covariates and predictors of political corruption in Malaysia using machine learning methods. Using the "Varieties of Democracy" dataset, hosted by the V-Dem Institute in Gothenburg, Sweden, it is hoped that variable selection, decision trees and random forests can be designed to hone in on the key indexes in the Vdem Dataset, to provide a suitable framework for future data analysis on political corruption. Variables that are identified will prove useful in the prediction of corruption trends not only within Malaysia, but with the additional objective of creating a generalisable approach for other nations. The initial country based analysis will cover the time period between 1980 to 2020, as within this period the country was ruled by three individuals, PM Tun Mahathir Mohamad (who was in office as Prime Minister twice, as Malaysia's 4th and 7th national leader), PM Abdullah Badawi, PM Najib Razak and PM Muhyiddin Yassin. This timeframe was chosen due to the upward trend of political corruption in Malaysia, and should allow the machine learning algorithms employed to identify important variables that can also be applied to other nations that have experienced growth in political corruption. After these predictors have been identified, a brief application will be applied to a neighbouring nation in Southeast Asia, the Philippines, with the expectation that the models will be able to achieve a similar (if somewhat) lower capacity to explain the variability of corruption in another nation.  

This study is primarily an exploratory data analysis, to identify key indicators using an ensemble of machine learning techniques. These identified variables will then be compared to qualitative sources in the discussion section, to confirm their importance to studies on political corruption, or even to take note of unexpected variables that were chosen. Limited testing of the identified variables through a linear model will be applied to the Philippines in later sections, to explore the external validity of the approach. The groundwork of this approach will essentially lay the foundation for future modeling studies that will incorporate regional data, perhaps to identify trends in political corruption in all of the nations of Southeast Asia, and if the generalisability of the approach is sound at this point, could be used to develop a more complex model on political corruption that includes other regions. 

## Research Questions

What covariates can be identified as key predictors of the changes witnessed in Malaysia's political corruption index in the past four decades? Do the identified covariates align with theories on corruption and good governance within the greater literature? Are these predictors also significant in modelling corruption when applied to a regional neighbour, the Philippines?"   

## Literature Review

Following a brief literature review, it seems that academic resources related to analyses of corruption in Malaysia remain largely limited to studies of institutional level corruption (and these studies are often related to specific public sector agencies), and with few sources conducting data based analyses on corruption. In a study by Kapeli & Mohamed for instance, they consider the public sector to be the largest concern in terms of corruption, and argued that low political will, ignorance of the causes of corruption, duplication of anti-corruption initiatives, and low public support for battling corruption as=re the primary reasons that institutional corruption has worsened in Malaysia in the past two decades [@kapeli2019battling, pg 552-554]. 

In an earlier article by the same authors, Kapeli & Mohamed note that the continued worsening of corruption perceptions in Malaysia in the past decade were also a result of the continued failure of anti-corruption oversight institutions, such as the Malaysian Anti-Corruption Comission (MACC), which had been established in 1997 originally as the Anti-Corruption Agency [@kapeli2015insight, pg 534]. One of their primary arguments in the article is centred on the fact that although Malaysia has actually developed extensive and elaborate frameworks and strategies to tackle corruption (through agencies such as the MACC), these systems are obstructed by inherent defects in the country’s overarching political systems, cultures and institutions [@kapeli2015insight, pg 528]

This assessment is supported by Hashim, who argues that efforts to tackle institutional corruption in Malaysia have failed due to a development of a "double standard" in the value system of key public officials in Malaysia, where key national leaders "allow [and] tolerate pilfering and pillaging by public officials, to persist." [@hashim2017development, pg 560]. This is essentially a result of top-down leadership in the country failing to address the development of an increasingly corrupt culture within government institutions, as they themselves develop intricate webs of public and private sector actors that conduct corrupt activities, making any efforts to detect and investigate such activities difficult and tedious [@hashim2017development, pg 559].

Unfortunately, the recent 2018 election in which Barisan Nasional was removed from power has not generated a large body of academic study on the subject, though it is likely we will see more research on the matter in the coming years. However, in an article by Ben Bland from 2019, he succintly describes the ousting of former PM Najib Razak's led government at the hands of Pakatan Harapan and Tun Dr. Mahathir Mohamad: 

"The coalition rode a wave of anger against the perceived corruption of the Najib government. It represents one of the most dramatic, bloodless rejections of authoritarian rule in recent global politics, with the United Malays National Organisation — in power for more than 60 years — ousted at the ballot box. Even more remarkable is the fact that this reformist revolution was led by former strongman Mahathir Mohamad, prime minister from 1981 to 2003, in alliance with his protégé-turned-nemesis Anwar Ibrahim." [@BenBland2018Mtot]

Bland goes on to note that while repression, electoral manipulation and allegations of corruption reached their zenith under PM Najib Razak, that critics of the newly elected government remain wary, and have concerns that a government under Tun Dr. Mahathir may return to a state of reduced transparency and a return of public sector corruption [@BenBland2018Mtot, pg 2]. It is apparent that while Mahathir has appointed some leading civil society advocates and technocrats to economic and institutional reform committees, Bland notes that certain appointments have been regarded with suspicion, such as Daim Zainuddin as a key economic adviser given that many Malaysians doubt that the former finance minister (Zainuddin) is likely to be a genuine reformer [@BenBland2018Mtot, pg 2].

In the larger context of corruption literature in the field of political science, there are many determinants of corruption that have been identified by various sources. In an analysis of four member states of the Association of Southeast Asian Nations (ASEAN), Sari et al [@sari2020coercive, pg 11-13] highlight the differing capacities for corruption disclosure practices in Indonesia, Thailand, the Philippines and Vietnam as key determinants for corruption levels domestically. They also note that structural organisation of domestic anti-corruption agencies has a major role to play in allowing an effective framework to fight political and other forms of corruption, and hence a key indicator of corruption levels. Unfortunately, a measure with this level of specificity does not exist in the Vdem Dataset, but can be accounted for by one of the measures for rule of law, especially those such as executive and judicial accountability.

On the issue of rule of law, and its many principles and components being a key part of reducing political corruption levels, it has been argued that in the case of Malaysia, there exists extensive legal frameworks for monitoring and combating corruption, but enforcement is completely lacking [@Siddiquee2005, pg 126]. This issue of enforcement is a key part of determining the true extent that rule of law is being upheld, as any country can have extensive anti-corruption mechanisms, but completely falls short on implementation and enforcement. In a multiple regression analysis by Shim & Eom, rule of law was found to be a significant predictor for corruption in three of four models they had developed, and especially in the sense of law enforcement on corruption (for which they used an overarching rule of law index as a proxy) [@shim2008government, pg 310]. They also argued that, aside from law enforcement, two key components that were often cited as key anti-corruption measures were establishing professionalism amongst civil servants, and enhancing bureaucratic quality overall, which they found to be significant as determinants of corruption in their models as well. In light of these theories, it is expected that a number of indicators covering transparency, accountability, law enforcement and other key parts of good governance will be identified by later variable selection.



## Data Source and Methodology

The data is sourced from the V-Dem Institute [@VDemData, 2020], which is a private research institute based in Gothenburg, Sweden. The dataset that the institute maintains is "Varieties of Democracy", an extensive dataset featuring policy data, measures of democratisation and other key covariates that relate to the political structure and institutions of the world's nations, with data reaching back over a century. As they state, the Varieties of Democracy dataset is a "multidimensional and disaggregated dataset that reflects the complexity of the concept of democracy as a system of rule that goes beyond the simple presence of elections. The V-Dem project distinguishes between five high-level principles of democracy: electoral, liberal, participatory, deliberative, and egalitarian, and collects data to measure these principles." [@VDemData, 2020]

The V-Dem dataset is a truly extensive dataset, featuring 4108 variables, with over 27,000 observations when taking into account the historical data of all observed nations. In the case of Malaysia, data goes back as far 1900, when the nation was officially administered as the Federated Malay States, a protectorate of Great Britain. The recorded data for Malaysia in the period of interest (between 1980 to 2020) is also extensive and appears to be frequently updated between years for variables of interest. While there is some missing data, these gaps are limited to variables that are unrelated to the question of political corruption, and will be eschewed for the purposes of this study. 

The methodological strategy of this study is to first clean and filter the data as much as possible, to reduce the risk of having variables with high collinearity being fed into the variable selection models. This is certainly a risk given that many of the 4000 over variables in Vdem are recoded and rescaled versions of primary versions of the data. Furthermore, a filter will be applied to remove variables that lack high levels of differentiation, that is, where observations have the same recorded value for intervals longer than three years. This is a necessary step to remove variables in which data collection by the Vdem Institute and their associates is more limited, and a static value is given for a period of 5 years for instance. Although some predictors of interest may be lost due to this procedure, it also serves the purpose of removing variables with single factor labels that would interfere with variable selection algorithms, while also keeping only variables with higher levels of differentiation in the data allowing the algorithms to learn and fit trends more accurately.

Variable selection will be conducted first, which will take the remaining three hundred variables (after filters and data cleaning), and locate the six most important. Variable selection will be accompanied with decision trees to minimise the number of variables further, though in balance with maintaining a high amount of the variation in the outcome variable explained by these predictors. Following this, a linear model will be fitted to the data using the selected variables, to confirm the statistical significance of the selection. The next approach will involve Random Forest algorithms, which will automatically choose variables by their importance in explaining the highest proportion of variation in the outcome, and can do so with the breadth of all the variables included. The most important of these variables will then be used to fit a linear model, to identify 

As a final step, both sets of predictors will be used to conduct a brief test of generalisability on a regional neighbour of Malaysia, the Philippines. The Philippines has been chosen not on the basis of having similar development levels, which the two countries do not have in common, but because of the widespread allegations of political corruption that exists in the country. Sari et al compared the levels of corruption disclosure practices (essentially public disclosure and reporting on incidents of corruption) in four nations within ASEAN, and found the Philippines to have the most pervasive level of corruption in its private sector, while also having the weakest anti-corruption agencies of the four. For this reason, the Philippines should prove to be an ideal comparative subset of the data in terms of high levels of corruption. 


### Data Cleaning

```{r}
# Read in Vdem Dataset
vdem_data <- readRDS("Data/V-Dem-CY-Full+Others-v11.1.rds")
```

```{r}
# Removal of NAs and filtering year to 1980 onwards
malaysia_data <- vdem_data %>%
  filter(country_name == "Malaysia")

coredata <- malaysia_data %>% 
  filter(year >= "1980")

coredata <- coredata %>%
  select_if(~ !any(is.na(.)))

unique_data <- sapply(lapply(coredata, unique), length)
unique_prune <- coredata[ , unique_data >= 3]
```


```{r}
# Regex to be used to filter additional versions of variables included in Vdem
# Ordinal scale, mean, standard deviation measures etc. 

versions <- "(_osp|_ord|_codelow|codehigh|_sd|_mean|_nr|_3C|_4C|_5C|e_|commnt)"

prune_ver <- grep(versions, colnames(unique_prune), value=TRUE, ignore.case =F)

clean_data <- unique_prune %>%
  select(-prune_ver)
```


### Variable Selection and Decision Tree


```{r}
# Subset Selection to identify significant variables. 
# v2xnp_regcorr removed due to high collinearity with v2x_corr
set.seed(333)
corrupt_subset <- regsubsets(v2x_corr ~. -v2xnp_regcorr, data = clean_data, method = "forward")
```


```{r fig.height=8, fig.width=10}
corsubset_rss <- subsets(corrupt_subset, statistic="rss", legend = TRUE, max.size = 6, main = "Corruption Subset: Residual Sum of Squares")
```

```{r fig.height=8, fig.width=10}
corsubset_cp <- subsets(corrupt_subset, statistic="cp", legend = TRUE, max.size = 6, main = "Corruption Subset: Mallow's Cp")
```

```{r fig.height=8, fig.width=10}
corsubset_bic <- subsets(corrupt_subset, statistic="bic", legend = TRUE, max.size = 6, main = "Corruption Subset: Bayesian Information Criteria")
```



```{r fig.height=8, fig.width=10}
corsubset_adjr2 <- subsets(corrupt_subset, statistic="adjr2", legend = TRUE, max.size = 6, main = "Corruption Subset: Adjusted R-squared")
```

Using the subsets function, the earlier regsubsets variable selection that narrowed down the number of variables from hundreds to just six can be easily translated to plots that also show the Mallows Cp, Bayesian Information Criteria, Adjusted R-squared and Residual Sum of Squares of the chosen subset. We find the optimal subset size, as determined by the previously mentioned criteria, is found to be six, and includes the following indicators (names were initially abbreviated by the function to ease computation): 

1. v2l_13 = Subnational election area less free and fair characteristics (C) (v2elsnlfc), Areas that are remote (difficult to reach by available transportation, for example). (0=No, 1=Yes) 

2. v2lgnl = Percentage of indirectly elected legislators upper chamber (A) (v2lginelup)

3. v2jcc = Judicial accountability (C) (v2juaccnt)

4. v2clrgs_3 = Stronger civil liberties characteristics (C) (v2clrgstch) = 3: Areas that are more economically developed. (0=No, 1=Yes)

5. v2svd = Domestic autonomy (C) (v2svdomaut)

6. v2x_x = Executive corruption index (D) (v2x_execorr)

These six variables can then be input into a decision tree, so that the algorithm can determine the smallest number of regressors that can achieve the highest accuracy in predicting variability in political corruption. 

```{r}
newtree <- tree(v2x_corr ~ v2elsnlfc_13 + v2lginelup + v2juaccnt + v2clrgstch_3 + v2svdomaut + v2x_execorr, data = clean_data)
```

```{r}
summary(newtree)
```

```{r}
plot(newtree, main = "Political Corruption Decision Tree: Malaysia")
text(newtree)

```

From the above decision tree summary and plot, we are given the three most important predictors from this method, that will be taken and applied to a linear model on corruption. It should be noted that the three variables identified are the executive corruption index (v2x_execorr), judicial accountability (v2juaccnt) and category 3 responses of the stronger civil liberties characteristics index (v2clrgstch). Only these three predictors will be needed, as the express purpose of utilising a decision tree was to minimise the number of predictors needed in explaining the largest proportion of variance. 

```{r}
set.seed(777)
treevar.fit <- lm(v2x_corr ~ v2x_execorr + v2juaccnt + v2clrgstch_3, data = clean_data)
```

```{r}
# Variance inflation factor for tree based method
vif(treevar.fit)
```

The variance inflation factor of the three selected variables is slightly high, though acceptable as they have not exceeded the value 3. Although these VIF values are by no means ideal, it has to be taken into account that many of the variables in the Vdem dataset are constructed from other indexes, making it difficult to sift through the hundreds of indicators included. Any variable with a VIF higher than 3 will be deemed unacceptable, and will be discarded through another filtration process, until a set of variables with acceptable VIF and Adjusted R-squared values has been reached.  This specification of a VIF value no greater than 3 will also be applied to the Random Forest approach. 

### Random Forest Approach

To take into account multicollinearity, a number of variance inflation factor tests were conducted on the variables that were highlighted to be of importance. The Rule of Law index has been removed due to its value being aggregated from a number of other indicators that can be seen as components of rule of law principles, such as judicial and executive accountability. Another set of examples would be the public sector and legislative corruption indexes, which were removeddue to their VIF values being over 40. The variables removed can be seen in the formula call of the corrupt_rf model below. As mentioned earlier, any variables with a VIF higher than 3 were removed to prevent multicollinearity issues. 

```{r}
set.seed(369)
corrupt_rf <- randomForest(v2x_corr ~. -v2x_rule -v2x_neopat -v2x_execorr -v2lgcrrpt -v2x_pubcorr -v2jucorrdc
                           -v2juhccomp-v2xnp_regcorr-v2juaccnt-v2jupack-v2exembez-v2clrspct-v2xcl_rol-v2jucomp
                           -v2x_jucon -v2mecrit -v2xnp_pres -v2clacjstm -v2xcl_prpty -v2clprptym -v2x_clpriv, 
                           data = clean_data)
```


```{r}
corrupt_rf
```

The random forest has incorporated all of the remaining predictors from the filtered list, and created a model that has arrived at a very low mean squared error of 0.00133, though with the over 280 predictors remaining a lot of noise has been introduced that has reduced its predictive capacity. Hence, we are given approximately 80% of the variance in political corruption being explained by the selection. Once the number of predictors has been reduced greatly, we should be able to construct a model with higher predictive accuracy.


```{r fig.height=8, fig.width=10}
varImpPlot(corrupt_rf, main = "Corrupt RF Variable Importance Plot")
```

From the above variable importance plot, the three most important predictors from this random forest model will be taken and applied to a linear model on corruption. It should be noted that the three variables that have the highest node purity identified are the Egalitarian component index (v2x_egal), executive bribery and corrupt exchanges (v2exbribe) and public sector corrupt exchanges (v2excrptps). Only these three will be selected, in the interest of choosing a small subset of similar size to that chosen by the tree method. 


```{r}
set.seed(0101)
forest.fit <- lm(v2x_corr ~ v2exbribe + v2x_egal + v2excrptps, data = clean_data)
```

```{r}
# Variance inflation factor for forest based method
vif(forest.fit)
```

In comparison to the tree method's model, the predictors chosen by the random forest method appear to have lower VIF values on average when fit to a linear model. Two of these predictors, the egalitarian index and executive bribery, have VIF values greater than 2 which is not ideal. 

Below can be found two summaries for the models that were fit using variables selected by either method. In the first, we see the random forest's model, which included executive bribery, egalitarian index and public sector corrupt exhanges as explanatory variables. We are given a multiple R-squared of 0.979 and an adjusted R-squared which is only slightly reduced, at 0.977. These values are only important in the context of this study in comparing the amount of variance explained by the competing predictors chosen by one of the two variable selection methods. It is also worth stating that we can observe a marked increase in variability explained (80% with the full set of predictors, compared to 97.9% with only the three most important) by the random forest approach by shrinking the number of predictors included. 

This summary is most valuable in the sense that it can give us an indication of the statistical significance of the chosen predictors. When we observe the t-values and p-values, we can surmise that all three are highly statistically significant, and that the overall model is also significant given the miniscule size of its p-value. As such, we can deduce that these three variables are indeed highly valuable predictors of political corruption in the context of Malaysia. 

```{r}
summary(forest.fit)
```



```{r}
summary(treevar.fit)
```


```{r}
# Random Forest model's residual plots
par(mfrow = c(2,2))
plot(forest.fit)
```


```{r}
# Decision Tree model's residual plots
par(mfrow = c(2,2))
plot(treevar.fit)
```


In comparing the residuals of the two plots, it would seem that the variables chosen by the random forest approach have presented themselves in a somewhat skewed pattern. But the assumption of linearity seems to hold, as there is no apparent presence of a pattern in the residuals, while the qqplot appears relatively normal, although the presence of heavy tails is noted.  The residuals vs leverage plot reveals a handful of outliers, however it is less of a concern compared to the decision tree approach. Furthermore, we can see that across the plots for the two models, that the outliers highlighted are in fact observations 38, 40 and 41, which are amongst the most recent years in which variability of the political corruption index in Malaysia has been great. 

In the decision tree method's customised model, we can see from the residual plots that the assumption of linearity appears to be at question, as the residuals are concentrated towards the right. The qqplot appears to indicate normality, though with heavy tails also putting the assumption at risk. The leverage plot also shows at least 3 points with high leverage, which would be a concern if we did not already know of the rapid decline in perceived corruption levels in those years. Overall, it seems that a linear model will suffice for the moment, though a larger sample of data (perhaps multiple countries and a longer timeframe) would be necessary to see if these assumptions hold. 




## Test of Generalisability

At this stage of the analysis, the variables selected by both the tree and forest based methods will be applied to the Philippines. To create a similar subset of the Vdem dataset for the Philippines, the same procedure will be applied for filtering and cleaning. 

```{r}
# Removal of NAs and filtering year to 1980 onwards, Country selection of the Philippines only
philippines_data <- vdem_data %>%
  filter(country_name == "Philippines")

testdata <- philippines_data %>% 
  filter(year >= "1980")

testdata <- testdata %>%
  select_if(~ !any(is.na(.)))

# No columns with fewer than 3 unique values will be included, to prevent modelling errors.
# unique_phil created as a vector to allow filtering
unique_data2 <- sapply(lapply(testdata, unique), length)
unique_phil <- testdata[ , unique_data2 >= 3]
```


```{r}
# Regex ("versions") will be applied, to filter different versions of key indices
# including ordinal, mean and others. 

versions2 <- "(_osp|_ord|_codelow|codehigh|_sd|_mean|_nr|_3C|_4C|_5C|e_|commnt)"

# Filter variables with non-unique/low variability values 
prune_phil <- grep(versions2, colnames(unique_phil), value=TRUE, ignore.case =F)

clean_data2 <- unique_phil %>%
  select(-prune_phil)
```

```{r}
# Linear Model using tree method's selected  variables
set.seed(121212)
tree.fitphil <- lm(v2x_corr ~ v2x_execorr + v2juaccnt + v2clrgstch_3, data = clean_data2)
```

```{r}
# Linear Model using random forest method's selected variables
set.seed(212121)
forest.fitphil <- lm(v2x_corr ~ v2exbribe + v2x_egal + v2excrptps, data = clean_data2)
```

```{r}

summary(tree.fitphil)
```

```{r}

summary(forest.fitphil)
```

We can immediately see that the variables selected by the decision tree method provide a model with a better fit, in comparison to the random forest selected variables. Both models can be considered statistically significant, given overall model p-values for both are very small. Howvever, if we compare the individual p-values of coefficients, we find that one of the predictors in the forest based model cannot be considered significant at the 0.05 level of significance, while another (public sector corruption, v2excrptps) is almost at the point of passing this significance threshold as well. 

In comparing the multiple and adjusted R-squared values, the predictors chosen by the tree based method definitively explain a greater proportion of the variability of political corruption in the Philippines, with at the very least 96.4% of variance in corruption explained, in comparison to the alternative model, which achieved an adjusted R-squared of 0.878, or 87.8% of the variance explained by included indicators. This is particularly good, considering that the predictors for executive corruption, judicial accountability and stronger civil liberties characteristics (label 3, meaning respect for civil liberties are stronger in areas that are more economically developed), could explain a similar level of variability of corruption in Malaysia (with the same model explaining 97.7% variance in Malaysian corruption). These results bode well for further application to future research that is inclusive of all nations within the greater ASEAN regional grouping. 

```{r}
# Decision Tree method based model, residual plots
par(mfrow = c(2,2))
plot(tree.fitphil)
```


```{r}
# Random Forest method based model, residual plots
par(mfrow = c(2,2))
plot(forest.fitphil)
```

The residuals plots of both models largely confirm the initial findings, in that it appears that the variables selected via decision tree have created a better fitting model as compared to the random forest approach. We see that OLS assumptions are on the whole held in the case of the tree based model, while in the other we can see irregularities in the residuals vs fitted plot and scale-location plot that would suggest non-linearity, while the qqplot also shows a number of observations as violating the assumption of normality. Overall, it seems that without applying non-linear terms to the model, it is more appropriate to go with the decision tree based linear model. 

## Discussion 

From the two separate methods utilised, we have arrived at a much smaller selection of predictor variables that can be used to explain a large proportion of the variance in political corruption in Malaysia. In the decision tree method, which utilised variable subset selection as an initial filter, three variables were identified as key regressors for political corruption, including the executive corruption index, judicial accountability and stronger civil liberties characteristics. In the random forest method, the three most important predictor variables that were selected include the egalitarian component index, executive bribery and corrupt exchanges and public sector corrupt exchanges. The two sets of variables are also listed below:


#### Decision Tree Method Selected Variables
1. Executive corruption index (v2x_execorr)
2. Judicial accountability (v2juaccnt)
3. Stronger civil liberties characteristics (v2clrgstch)

#### Random Forest Method Selected Variables
1. Egalitarian component index (v2x_egal)
2. Executive bribery and corrupt exchanges (v2exbribe)
3. Public sector corrupt exchanges (v2excrptps)

While both methods yielded models that could be considered statistically significant, and do in fact account for a large proportion of variance in corruption, the decision tree based method yielded a model that had greater success in its application to a different country setting, achieving approximately 10% higher explanatory power than the other. The results are somewhat unexpected in the case of the application of both models to the Philippines. In the Random Forest variable model, two of the predictors, public sector corruption and executive bribery were found to have lower statistical significance than some of the variables identified by the decision tree method. The coefficient for public sector corruption specifically, was found to have breached the 0.05 level of significance threshold, rendering its statistical reliability as questionable for further use. It seems that certain aspects of understanding growth of political corruption in the Malaysia-specific literature does not carry over to different contexts. Although public sector corruption was highlighted as a key component of the pervasive corruption experienced in Malaysia in the past few decade, as argued by Kapeli [@kapeli2019battling, pg 552 - 554], it appears that executive corruption is a better indicator in the case of the Philippines. 

The difference in the two variables comes down to the exact action undertaken, as the Vdem Institute distinguishes executive bribery as being an measure of how often a country's political executive (cabinet members, head of state etc.) "grant favors in exchange for bribes, kickbacks, or other material inducements" [@VDemCode, pg 112-113]. This variable is also on a positive ordinal scale from 0 to 4, where 0 is the response, "It is routine and expected", while 4 is the response, "It never, or hardly ever, happens". As such, we can see that as the value of the executive bribery variable goes up, we would expect a decrease in the political corruption index. Executive corruption on the other hand, has an added component, as while its definition includes the granting of favours in exchange for bribes and kickbacks, this measure also includes stealing, embezzlement and misappropriation of public funds and state resources for personal or family use. This added component is gives us clear indication that the executive corruption index is a far more complete metric. 

Executive corruption is a highly important determinant for the overall pervasiveness of political corruption within a country, as political executives tolerating corruption and bribery under their watch allows illicit behaviour and practices to develop as part of bureaucratic culture and structures, as was the case in Malaysia in the past two decades [@hashim2017development, pg 559]. This top-down effect of corrupt activities has wide-ranging implications, as by tolerating pervasive corruption at lower bureaucratic structures in the governmental hierarchy, there is also the propensity for corruption to become ingrained in private sector practices as well. This is in cosideration of the fact that civil servants have a duty not only to safeguard governance structures from political influence and corruption, but also to prevent private sector entities from exerting their own influence to secure special interests [@shim2008government, pg 300]. With no safeguard against these interests, corruption at the executive level can easily exacerbate overall political corruption by an inability to deter smaller scale bribery and favouritism between public and private sector entities. To see this variable chosen as a key determinant for corruption via variable selection is no surprise, and could very well be applicable to most countries facing pervasive corruption within political and bureaucratic structures. 

The selection of public sector corrupt exchanges by the random forest approach is very much in line with existing literature, as both Kapeli [@kapeli2019battling, pg 552 - 554] and Hashim [@hashim2017development, pg 559] highlighted the inadequacies of reporting and law enforcement against corrupt activities by public officials and civil servants. In the case of the latter, Hashim noted that the Malaysian Anti-Corruption Commission, or MACC, had a total of 3533 arrests for charges of corruption between the year 2011 and 2015. Of these individuals arrested, 1408 of them were public officials (which is nearly 40% of the total), demonstrating the high level of public sector corruption in recent years withi the country. 

In the broader literature, it has been posited that the "wage rate of civil servants relative to that in the private sector has an impact on the incidence of corruption. Relatively low wages in the public sector will make the benefit of a given bribe seem greater and the cost of losing the government job if the bribe is discovered seem less" [@elbahnasawy2012determinants, pg 314]. As such, it is unexpected that this indicator was not found to be more significant in the case of the Philippines, and may be a reflection of a more disciplined or well-enforced public sector relative to the downward trend in Malaysia.  

The indicator for levels of judicial accountability is somewhat puzzling, as the variable is a measure of survey responses on their perception of how often judges are held accountable within their countries when they are found to have comitted misconduct. 
Increasing units in this metric would imply greater rule of law and better governance oversight, but the positive sign of the coefficient implies that greater judicial accountability is associated with higher levels of corruption. This is in opposition to theory on the matter, as the 

The indicator for stronger civil liberties characteristics, which was specifically the perception that stronger respect for civil liberties by government officials in wealthier districts was found to have a negative relationship with political corruption. The implication of this finding seems to be that the public perceives political corruption to be lower, when there is evidence of rule of law being upheld, in the sense that individual civil liberties are strongest in wealthy areas. This could be as a response to the wealthy and other elites being left unmolested by unjust government interference and that their rights to freedom of action and speech are undenied. Some research on the relationship between civil liberties and corruption exists, with Roca & Alidedeoglu-Buchner identifying a significant statistical relationship between corruption and political rights, which they argue are closely related to civil liberties, as being inversely related to corruption perception [@roca2010corruption, pg 8-10]. In their paper however, the focus is on the maturity of a democracy and does not include any element of the economic wealth of citizens or districts being related to stronger civil liberties. Further investigation on this specific indicator is suggested, as its role as a determinant of corruption remains somewhat abstract at this juncture, given its specificity to the wealth of a region playing a role. It is however, a statistically significant predictor, given the p-value for the cofficient was 0.00106, and as such is worth maintaining for future research. 

## Conclusion 

To conclude, three predictors have been found to be the most important explanatory variables in explaining variation in political corruption in Malaysia. Of the two methods constructed, it is the variable selection and decision tree based method that has resulted in the most accurate predictive linear model. This is true of not only political corruption trends in Malaysia, but also in the Philippines, as was found during a brief generalisability test. Similar levels of variance were accounted for by the model in both countries, while remaining statistically significant at all conventional levels of significance. The identified variables include executive corruption, stronger civil liberties characteristics and judicial accountability, with these indicators falling largely in line with existing theories in political corruption literature. It is interesting to note that in the case of judicial accountability, the sign of the coefficient was unexpectedly positive, so there is a need for further investigation on the exact relationship between increasing levels of judicial accountability and political corruption. 

This investigation on key variables could prove invaluable in future research, not only on trends of political corruption in Malaysia, but also for its neighbours in the Association of Southeast Asian Nations, and perhaps a similar approach could be applied to other regions of the world. Further testing would be required, but considering the extensive nature of the Vdem Dataset, it is well within the means of other researchers to apply similar variable selection and predictive models as have been developed in this study, to their chosen context. 



## Bibliography


